---
title: "Mnemonic Similarity Task (MST) – Descriptive Statistics Analysis"
author: "Sourav Nath, Piyush Agarwal, Rajarshi Chatterjee"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
    theme: cosmo
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 8, fig.height = 5, dpi = 150)
```

# 1. Setup and Data Loading

```{r libraries}
# Load required libraries
library(ggplot2)
library(dplyr)
library(tidyr)
library(knitr)

# Clear workspace
rm(list = setdiff(ls(), lsf.str()))
```

## 1.1 Define Helper Functions

```{r helper-functions}
# Function to read all test files from a condition folder
read_test_files <- function(data_dir, condition_name) {
  # List all test CSV files
  test_files <- list.files(data_dir, pattern = "MST_test.*\\.csv$", full.names = TRUE)
  
  all_data <- data.frame()
  
  for (f in test_files) {
    # Extract participant ID from filename
    participant_id <- sub("^(\\d+)_MST_test.*", "\\1", basename(f))
    
    df <- read.csv(f, stringsAsFactors = FALSE)
    
    # Keep only trial rows (rows with actual image_path and response)
    # In test files, trial rows have non-empty key_resp_3.keys and image_path
    trial_cols <- c("thisN", "key_resp_3.keys", "key_resp_3.rt", 
                    "image_path", "position_of_stimuli")
    
    # Check which columns exist
    existing_cols <- trial_cols[trial_cols %in% names(df)]
    df_sub <- df[existing_cols]
    
    # Filter to actual trial rows
    df_sub <- df_sub[!is.na(df_sub$image_path) & df_sub$image_path != "" &
                     !is.na(df_sub$key_resp_3.keys) & df_sub$key_resp_3.keys != "", ]
    
    if (nrow(df_sub) > 0) {
      df_sub$participant <- participant_id
      df_sub$condition <- condition_name
      all_data <- rbind(all_data, df_sub)
    }
  }
  
  return(all_data)
}

# Function to read bins file
read_bins <- function(filepath) {
  bins <- read.delim(filepath, header = FALSE, sep = "\t", 
                     col.names = c("item_num", "bin"))
  return(bins)
}
```

## 1.2 Load Data from All Three Conditions

```{r load-data}
base_dir <- "MST_Data"

# Load test data from all conditions
item_only_data <- read_test_files(
  file.path(base_dir, "item_only", "item_only_data"), "Item Only")
task_only_data <- read_test_files(
  file.path(base_dir, "task_only", "task_only_data"), "Task Only")
both_data <- read_test_files(
  file.path(base_dir, "Both_item_task", "both_data"), "Both")

# Combine all data
all_test_data <- rbind(item_only_data, task_only_data, both_data)

# Convert response to factor
all_test_data$response <- factor(all_test_data$key_resp_3.keys, 
                                  levels = c("o", "s", "n"),
                                  labels = c("Old", "Similar", "New"))

# Convert reaction time to numeric
all_test_data$rt <- as.numeric(all_test_data$key_resp_3.rt)

# Classify stimulus type based on position_of_stimuli
# pre, mid, post = Lure (similar items at different positions)
# none = Foil (completely new)
# For targets: items shown at encoding that appear again exactly
# In this dataset: 'post' and 'pre' and 'mid' positions of lures
# The position_of_stimuli indicates the type:
#   - Items from Objects/Scenes folders with 'a' suffix shown at encoding = targets if shown again
#   - Items with 'b' suffix = lures (similar versions)
#   - Foils = brand new items from Foils folder

all_test_data$stimulus_type <- ifelse(
  grepl("^Foils", all_test_data$image_path), "Foil",
  ifelse(grepl("b\\.", all_test_data$image_path), "Lure", "Target")
)

all_test_data$stimulus_type <- factor(all_test_data$stimulus_type, 
                                      levels = c("Target", "Lure", "Foil"))

# Convert condition to factor
all_test_data$condition <- factor(all_test_data$condition, 
                                  levels = c("Item Only", "Task Only", "Both"))

cat("Total trials loaded:", nrow(all_test_data), "\n")
cat("Participants per condition:\n")
print(tapply(all_test_data$participant, all_test_data$condition, 
             function(x) length(unique(x))))
cat("\nTrials per stimulus type:\n")
print(table(all_test_data$stimulus_type))
```

## 1.3 Load Similarity Bins

```{r load-bins}
# Load bins for Set6 (Objects)
bins_set6 <- read_bins(file.path(base_dir, "item_only", "Set6 bins.txt"))
bins_set6$set <- "Objects"

# Load bins for SetScC (Scenes)
bins_scc <- read_bins(file.path(base_dir, "item_only", "SetScC bins.txt"))
bins_scc$set <- "Scenes"

# Combine bins
all_bins <- rbind(bins_set6, bins_scc)
cat("Bins distribution:\n")
print(table(all_bins$bin))
```

# 2. Compute Per-Participant Summary Metrics

```{r compute-metrics}
# For each participant, compute key metrics
participant_metrics <- all_test_data %>%
  group_by(participant, condition) %>%
  summarise(
    # Total trials by stimulus type
    n_targets = sum(stimulus_type == "Target", na.rm = TRUE),
    n_lures = sum(stimulus_type == "Lure", na.rm = TRUE),
    n_foils = sum(stimulus_type == "Foil", na.rm = TRUE),
    
    # Hit Rate: P("Old" | Target)
    hit_rate = sum(stimulus_type == "Target" & response == "Old", na.rm = TRUE) / 
               sum(stimulus_type == "Target", na.rm = TRUE),
    
    # False Alarm Rate: P("Old" | Foil)
    fa_rate = sum(stimulus_type == "Foil" & response == "Old", na.rm = TRUE) / 
              sum(stimulus_type == "Foil", na.rm = TRUE),
    
    # Correct Rejection Rate: P("New" | Foil)
    cr_rate = sum(stimulus_type == "Foil" & response == "New", na.rm = TRUE) / 
              sum(stimulus_type == "Foil", na.rm = TRUE),
    
    # Lure Discrimination Index (LDI): P("Similar" | Lure) - P("Similar" | Foil)
    p_sim_lure = sum(stimulus_type == "Lure" & response == "Similar", na.rm = TRUE) / 
                 sum(stimulus_type == "Lure", na.rm = TRUE),
    p_sim_foil = sum(stimulus_type == "Foil" & response == "Similar", na.rm = TRUE) / 
                 sum(stimulus_type == "Foil", na.rm = TRUE),
    LDI = p_sim_lure - p_sim_foil,
    
    # Recognition Memory (REC): P("Old" | Target) - P("Old" | Foil)
    REC = hit_rate - fa_rate,
    
    # Mean Reaction Times
    rt_targets = mean(rt[stimulus_type == "Target"], na.rm = TRUE),
    rt_lures = mean(rt[stimulus_type == "Lure"], na.rm = TRUE),
    rt_foils = mean(rt[stimulus_type == "Foil"], na.rm = TRUE),
    rt_overall = mean(rt, na.rm = TRUE),
    
    .groups = "drop"
  )

cat("Participant metrics computed for", nrow(participant_metrics), "participants\n")
```

## 2.1 Summary Statistics Table

```{r summary-table}
# Compute group-level summary statistics
summary_stats <- participant_metrics %>%
  group_by(condition) %>%
  summarise(
    N = n(),
    `Hit Rate (M)` = round(mean(hit_rate, na.rm = TRUE), 3),
    `Hit Rate (SD)` = round(sd(hit_rate, na.rm = TRUE), 3),
    `FA Rate (M)` = round(mean(fa_rate, na.rm = TRUE), 3),
    `FA Rate (SD)` = round(sd(fa_rate, na.rm = TRUE), 3),
    `LDI (M)` = round(mean(LDI, na.rm = TRUE), 3),
    `LDI (SD)` = round(sd(LDI, na.rm = TRUE), 3),
    `REC (M)` = round(mean(REC, na.rm = TRUE), 3),
    `REC (SD)` = round(sd(REC, na.rm = TRUE), 3),
    `CR Rate (M)` = round(mean(cr_rate, na.rm = TRUE), 3),
    `CR Rate (SD)` = round(sd(cr_rate, na.rm = TRUE), 3),
    .groups = "drop"
  )

kable(summary_stats, caption = "Table 1: Summary Statistics by Condition (Mean ± SD)")
```

# 3. Response Proportion Analysis

```{r response-proportions}
# Compute response proportions for each stimulus type × condition
response_props <- all_test_data %>%
  group_by(condition, stimulus_type, response) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(condition, stimulus_type) %>%
  mutate(proportion = count / sum(count)) %>%
  ungroup()

# Display as table
response_table <- response_props %>%
  select(condition, stimulus_type, response, proportion) %>%
  pivot_wider(names_from = response, values_from = proportion, values_fill = 0) %>%
  mutate(across(where(is.numeric), ~ round(., 3)))

kable(response_table, caption = "Table 2: Response Proportions by Stimulus Type and Condition")
```

## 3.1 Figure 1: Response Proportions by Stimulus Type

```{r fig1-response-proportions, fig.width=10, fig.height=6}
ggplot(response_props, aes(x = stimulus_type, y = proportion, fill = response)) +
  geom_bar(stat = "identity", position = "dodge", color = "black", size = 0.3) +
  facet_wrap(~ condition) +
  scale_fill_manual(values = c("Old" = "#E74C3C", "Similar" = "#3498DB", "New" = "#2ECC71"),
                    name = "Response") +
  labs(title = "Figure 1: Response Proportions by Stimulus Type Across Conditions",
       x = "Stimulus Type", y = "Proportion of Responses") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", size = 14),
        strip.text = element_text(face = "bold", size = 12),
        legend.position = "bottom")

ggsave("figures/fig1_response_proportions.png", width = 10, height = 6, dpi = 300)
```

# 4. Key Metrics Comparison Across Conditions

## 4.1 Figure 2: LDI and REC Across Conditions

```{r fig2-ldi-rec, fig.width=10, fig.height=5}
# Prepare data for plotting
metrics_long <- participant_metrics %>%
  select(participant, condition, LDI, REC, hit_rate, fa_rate, cr_rate) %>%
  pivot_longer(cols = c(LDI, REC, hit_rate, fa_rate, cr_rate),
               names_to = "metric", values_to = "value")

# LDI and REC boxplots
ldi_rec_data <- metrics_long %>%
  filter(metric %in% c("LDI", "REC"))

ggplot(ldi_rec_data, aes(x = condition, y = value, fill = condition)) +
  geom_boxplot(alpha = 0.7, outlier.shape = 21) +
  geom_jitter(width = 0.15, alpha = 0.3, size = 1.5) +
  facet_wrap(~ metric, scales = "free_y") +
  scale_fill_manual(values = c("Item Only" = "#9B59B6", 
                                "Task Only" = "#E67E22", 
                                "Both" = "#1ABC9C")) +
  labs(title = "Figure 2: Lure Discrimination Index (LDI) and Recognition (REC) by Condition",
       x = "Condition", y = "Score") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", size = 14),
        strip.text = element_text(face = "bold", size = 12),
        legend.position = "none")

ggsave("figures/fig2_ldi_rec_boxplot.png", width = 10, height = 5, dpi = 300)
```

## 4.2 Figure 3: Hit Rate and False Alarm Rate

```{r fig3-hit-fa, fig.width=10, fig.height=5}
hit_fa_data <- metrics_long %>%
  filter(metric %in% c("hit_rate", "fa_rate")) %>%
  mutate(metric = recode(metric, 
                         "hit_rate" = "Hit Rate", 
                         "fa_rate" = "False Alarm Rate"))

ggplot(hit_fa_data, aes(x = condition, y = value, fill = condition)) +
  geom_boxplot(alpha = 0.7, outlier.shape = 21) +
  geom_jitter(width = 0.15, alpha = 0.3, size = 1.5) +
  facet_wrap(~ metric) +
  scale_fill_manual(values = c("Item Only" = "#9B59B6", 
                                "Task Only" = "#E67E22", 
                                "Both" = "#1ABC9C")) +
  labs(title = "Figure 3: Hit Rate and False Alarm Rate by Condition",
       x = "Condition", y = "Rate") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", size = 14),
        strip.text = element_text(face = "bold", size = 12),
        legend.position = "none")

ggsave("figures/fig3_hit_fa_boxplot.png", width = 10, height = 5, dpi = 300)
```

# 5. Distribution Analysis

## 5.1 Figure 4: Distribution of LDI Scores

```{r fig4-ldi-distribution, fig.width=10, fig.height=5}
# Histogram with overlaid normal curve
ggplot(participant_metrics, aes(x = LDI)) +
  geom_histogram(aes(y = after_stat(density)), bins = 20, 
                 fill = "#3498DB", color = "white", alpha = 0.7) +
  geom_density(color = "#E74C3C", linewidth = 1.2) +
  facet_wrap(~ condition) +
  labs(title = "Figure 4: Distribution of Lure Discrimination Index (LDI) by Condition",
       x = "LDI Score", y = "Density") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", size = 14),
        strip.text = element_text(face = "bold", size = 12))

ggsave("figures/fig4_ldi_distribution.png", width = 10, height = 5, dpi = 300)
```

## 5.2 Normal Distribution and t-Distribution Fitting

```{r distribution-fitting}
# For each condition, compare LDI distribution to normal and t-distribution
cat("=== Shapiro-Wilk Normality Tests for LDI ===\n\n")

for (cond in levels(participant_metrics$condition)) {
  ldi_vals <- participant_metrics$LDI[participant_metrics$condition == cond]
  ldi_vals <- ldi_vals[!is.na(ldi_vals)]
  
  cat(paste0("Condition: ", cond, "\n"))
  cat(paste0("  N = ", length(ldi_vals), "\n"))
  cat(paste0("  Mean = ", round(mean(ldi_vals), 4), "\n"))
  cat(paste0("  SD = ", round(sd(ldi_vals), 4), "\n"))
  cat(paste0("  Skewness = ", round(moments::skewness(ldi_vals), 4), "\n"))
  cat(paste0("  Kurtosis = ", round(moments::kurtosis(ldi_vals), 4), "\n"))
  
  sw_test <- shapiro.test(ldi_vals)
  cat(paste0("  Shapiro-Wilk W = ", round(sw_test$statistic, 4), 
             ", p = ", round(sw_test$p.value, 4), "\n\n"))
}
```

```{r fig5-normal-t-comparison, fig.width=10, fig.height=5}
# Compare Normal vs t-distribution fit for LDI scores
plot_data <- data.frame()

for (cond in levels(participant_metrics$condition)) {
  ldi_vals <- participant_metrics$LDI[participant_metrics$condition == cond]
  ldi_vals <- ldi_vals[!is.na(ldi_vals)]
  
  m <- mean(ldi_vals)
  s <- sd(ldi_vals)
  n <- length(ldi_vals)
  
  x_seq <- seq(min(ldi_vals) - 0.1, max(ldi_vals) + 0.1, length.out = 200)
  
  # Normal density
  normal_density <- dnorm(x_seq, mean = m, sd = s)
  
  # t-distribution density (scaled) with n-1 degrees of freedom
  t_density <- dt((x_seq - m) / s, df = n - 1) / s
  
  temp <- data.frame(
    x = rep(x_seq, 2),
    density = c(normal_density, t_density),
    distribution = rep(c("Normal", "t-Distribution"), each = length(x_seq)),
    condition = cond
  )
  plot_data <- rbind(plot_data, temp)
}

ggplot() +
  geom_histogram(data = participant_metrics, 
                 aes(x = LDI, y = after_stat(density)), 
                 bins = 15, fill = "grey80", color = "white", alpha = 0.6) +
  geom_line(data = plot_data, 
            aes(x = x, y = density, color = distribution, linetype = distribution),
            linewidth = 1.1) +
  facet_wrap(~ condition) +
  scale_color_manual(values = c("Normal" = "#E74C3C", "t-Distribution" = "#2980B9"),
                     name = "Distribution") +
  scale_linetype_manual(values = c("Normal" = "solid", "t-Distribution" = "dashed"),
                        name = "Distribution") +
  labs(title = "Figure 5: LDI Distribution with Normal and t-Distribution Overlay",
       x = "LDI Score", y = "Density") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", size = 14),
        strip.text = element_text(face = "bold", size = 12),
        legend.position = "bottom")

ggsave("figures/fig5_normal_t_comparison.png", width = 10, height = 5, dpi = 300)
```

## 5.3 One-Sample t-Tests (Descriptive)

```{r t-tests}
# Test whether LDI is significantly different from 0 (chance) for each condition
cat("=== One-Sample t-Tests: Is LDI significantly > 0? ===\n\n")

for (cond in levels(participant_metrics$condition)) {
  ldi_vals <- participant_metrics$LDI[participant_metrics$condition == cond]
  ldi_vals <- ldi_vals[!is.na(ldi_vals)]
  
  t_result <- t.test(ldi_vals, mu = 0, alternative = "greater")
  
  cat(paste0("Condition: ", cond, "\n"))
  cat(paste0("  Mean LDI = ", round(t_result$estimate, 4), "\n"))
  cat(paste0("  t(", t_result$parameter, ") = ", round(t_result$statistic, 4), "\n"))
  cat(paste0("  p-value = ", format(t_result$p.value, scientific = TRUE, digits = 4), "\n"))
  cat(paste0("  95% CI: [", round(t_result$conf.int[1], 4), ", Inf]\n"))
  
  # Cohen's d
  cohens_d <- mean(ldi_vals) / sd(ldi_vals)
  cat(paste0("  Cohen's d = ", round(cohens_d, 4), "\n\n"))
}

# Same for REC
cat("=== One-Sample t-Tests: Is REC significantly > 0? ===\n\n")

for (cond in levels(participant_metrics$condition)) {
  rec_vals <- participant_metrics$REC[participant_metrics$condition == cond]
  rec_vals <- rec_vals[!is.na(rec_vals)]
  
  t_result <- t.test(rec_vals, mu = 0, alternative = "greater")
  
  cat(paste0("Condition: ", cond, "\n"))
  cat(paste0("  Mean REC = ", round(t_result$estimate, 4), "\n"))
  cat(paste0("  t(", t_result$parameter, ") = ", round(t_result$statistic, 4), "\n"))
  cat(paste0("  p-value = ", format(t_result$p.value, scientific = TRUE, digits = 4), "\n"))
  cat(paste0("  95% CI: [", round(t_result$conf.int[1], 4), ", Inf]\n\n"))
}
```

# 6. Reaction Time Analysis

```{r rt-summary}
# RT summary statistics
rt_summary <- all_test_data %>%
  filter(!is.na(rt) & rt > 0) %>%
  group_by(condition, stimulus_type) %>%
  summarise(
    Mean_RT = round(mean(rt, na.rm = TRUE), 3),
    SD_RT = round(sd(rt, na.rm = TRUE), 3),
    Median_RT = round(median(rt, na.rm = TRUE), 3),
    N = n(),
    .groups = "drop"
  )

kable(rt_summary, caption = "Table 3: Reaction Time (seconds) by Condition and Stimulus Type")
```

## 6.1 Figure 6: Reaction Times by Stimulus Type

```{r fig6-rt, fig.width=10, fig.height=5}
rt_data <- all_test_data %>%
  filter(!is.na(rt) & rt > 0 & rt < quantile(rt, 0.99, na.rm = TRUE))

ggplot(rt_data, aes(x = stimulus_type, y = rt, fill = stimulus_type)) +
  geom_violin(alpha = 0.6, trim = TRUE) +
  geom_boxplot(width = 0.15, alpha = 0.8, outlier.size = 0.5) +
  facet_wrap(~ condition) +
  scale_fill_manual(values = c("Target" = "#E74C3C", 
                                "Lure" = "#F39C12", 
                                "Foil" = "#27AE60")) +
  labs(title = "Figure 6: Reaction Time Distributions by Stimulus Type and Condition",
       x = "Stimulus Type", y = "Reaction Time (seconds)") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", size = 14),
        strip.text = element_text(face = "bold", size = 12),
        legend.position = "none")

ggsave("figures/fig6_rt_violin.png", width = 10, height = 5, dpi = 300)
```

## 6.2 Figure 7: Mean RT by Response Type

```{r fig7-rt-response, fig.width=10, fig.height=5}
rt_by_response <- all_test_data %>%
  filter(!is.na(rt) & rt > 0 & !is.na(response)) %>%
  group_by(condition, response) %>%
  summarise(
    mean_rt = mean(rt, na.rm = TRUE),
    se_rt = sd(rt, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

ggplot(rt_by_response, aes(x = response, y = mean_rt, fill = condition)) +
  geom_bar(stat = "identity", position = "dodge", color = "black", size = 0.3) +
  geom_errorbar(aes(ymin = mean_rt - se_rt, ymax = mean_rt + se_rt),
                position = position_dodge(0.9), width = 0.2) +
  scale_fill_manual(values = c("Item Only" = "#9B59B6", 
                                "Task Only" = "#E67E22", 
                                "Both" = "#1ABC9C")) +
  labs(title = "Figure 7: Mean Reaction Time by Response Type Across Conditions",
       x = "Response", y = "Mean RT (seconds)",
       fill = "Condition") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", size = 14),
        legend.position = "bottom")

ggsave("figures/fig7_rt_by_response.png", width = 10, height = 5, dpi = 300)
```

# 7. Similarity Bin Analysis

```{r bin-analysis}
# Extract item number from image path for lure items
# Lure items have 'b' suffix: e.g., Objects\039b.jpg -> item 39
lure_data <- all_test_data %>%
  filter(stimulus_type == "Lure") %>%
  mutate(
    # Extract the number from the image path
    item_num_str = sub(".*[/\\\\](\\d+)b\\.jpg$", "\\1", image_path),
    item_num = as.integer(item_num_str),
    # Determine if object or scene
    stim_set = ifelse(grepl("^Objects", image_path), "Objects", "Scenes")
  ) %>%
  filter(!is.na(item_num))

# Merge with bins
# For Objects: use bins_set6, for Scenes: use bins_scc  
lure_with_bins <- lure_data %>%
  left_join(
    rbind(
      bins_set6 %>% mutate(stim_set = "Objects"),
      bins_scc %>% mutate(stim_set = "Scenes")
    ),
    by = c("item_num" = "item_num", "stim_set" = "stim_set")
  ) %>%
  filter(!is.na(bin))

# Compute LDI-like metric by bin  
bin_performance <- lure_with_bins %>%
  group_by(condition, bin) %>%
  summarise(
    n_trials = n(),
    p_old = mean(response == "Old", na.rm = TRUE),
    p_similar = mean(response == "Similar", na.rm = TRUE),
    p_new = mean(response == "New", na.rm = TRUE),
    mean_rt = mean(rt, na.rm = TRUE),
    .groups = "drop"
  )

# Table
kable(bin_performance %>% mutate(across(where(is.numeric), ~ round(., 3))),
      caption = "Table 4: Lure Performance by Similarity Bin (1=most similar, 5=least similar)")
```

## 7.1 Figure 8: Lure Performance by Similarity Bin

```{r fig8-bin-performance, fig.width=10, fig.height=6}
bin_props <- lure_with_bins %>%
  group_by(condition, bin, response) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(condition, bin) %>%
  mutate(proportion = count / sum(count)) %>%
  ungroup() %>%
  mutate(bin = factor(bin, labels = paste("Bin", 1:5)))

ggplot(bin_props, aes(x = bin, y = proportion, fill = response)) +
  geom_bar(stat = "identity", position = "stack", color = "black", size = 0.2) +
  facet_wrap(~ condition) +
  scale_fill_manual(values = c("Old" = "#E74C3C", "Similar" = "#3498DB", "New" = "#2ECC71"),
                    name = "Response") +
  labs(title = "Figure 8: Response Proportions for Lures by Similarity Bin",
       subtitle = "Bin 1 = most similar to target, Bin 5 = least similar",
       x = "Similarity Bin", y = "Proportion of Responses") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", size = 14),
        plot.subtitle = element_text(size = 11, color = "grey40"),
        strip.text = element_text(face = "bold", size = 12),
        legend.position = "bottom")

ggsave("figures/fig8_bin_performance.png", width = 10, height = 6, dpi = 300)
```

# 8. Lure Position Analysis (pre, mid, post)

```{r pos-analysis}
# Filter for lure items that have a specific position (pre, mid, post)
lure_pos_data <- all_test_data %>%
  filter(stimulus_type == "Lure" & position_of_stimuli %in% c("pre", "mid", "post"))

# Compute performance by position
pos_performance <- lure_pos_data %>%
  group_by(condition, position_of_stimuli) %>%
  summarise(
    n_trials = n(),
    p_old = mean(response == "Old", na.rm = TRUE),
    p_similar = mean(response == "Similar", na.rm = TRUE),
    p_new = mean(response == "New", na.rm = TRUE),
    mean_rt = mean(rt, na.rm = TRUE),
    .groups = "drop"
  )

# Table
kable(pos_performance %>% mutate(across(where(is.numeric), ~ round(., 3))),
      caption = "Table 5: Lure Performance by Position of Stimuli (pre, mid, post)")
```

## 8.1 Figure 10: Lure Performance by Position

```{r fig10-pos-performance, fig.width=10, fig.height=6}
pos_props <- lure_pos_data %>%
  group_by(condition, position_of_stimuli, response) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(condition, position_of_stimuli) %>%
  mutate(proportion = count / sum(count)) %>%
  ungroup() %>%
  mutate(position_of_stimuli = factor(position_of_stimuli, levels = c("pre", "mid", "post")))

ggplot(pos_props, aes(x = position_of_stimuli, y = proportion, fill = response)) +
  geom_bar(stat = "identity", position = "stack", color = "black", size = 0.2) +
  facet_wrap(~ condition) +
  scale_fill_manual(values = c("Old" = "#E74C3C", "Similar" = "#3498DB", "New" = "#2ECC71"),
                    name = "Response") +
  labs(title = "Figure 10: Response Proportions for Lures by Position",
       x = "Position of Stimuli", y = "Proportion of Responses") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", size = 14),
        strip.text = element_text(face = "bold", size = 12),
        legend.position = "bottom")

ggsave("figures/fig10_pos_performance.png", width = 10, height = 6, dpi = 300)
```

# 9. Additional Descriptive Statistics

## 9.1 Overall Accuracy

```{r overall-accuracy}
# Compute overall accuracy: correct responses for each stimulus type
# Target → "Old", Lure → "Similar", Foil → "New"
all_test_data$correct <- with(all_test_data, 
  (stimulus_type == "Target" & response == "Old") |
  (stimulus_type == "Lure" & response == "Similar") |
  (stimulus_type == "Foil" & response == "New")
)

accuracy_by_condition <- all_test_data %>%
  group_by(condition) %>%
  summarise(
    overall_accuracy = round(mean(correct, na.rm = TRUE), 3),
    target_accuracy = round(mean(correct[stimulus_type == "Target"], na.rm = TRUE), 3),
    lure_accuracy = round(mean(correct[stimulus_type == "Lure"], na.rm = TRUE), 3),
    foil_accuracy = round(mean(correct[stimulus_type == "Foil"], na.rm = TRUE), 3),
    .groups = "drop"
  )

kable(accuracy_by_condition, 
      caption = "Table 6: Accuracy Rates by Condition",
      col.names = c("Condition", "Overall", "Target", "Lure", "Foil"))
```

## 9.2 Participant-Level Variability

```{r fig11-participant-variability, fig.width=10, fig.height=5}
# Scatter plot: Hit Rate vs LDI colored by condition
ggplot(participant_metrics, aes(x = hit_rate, y = LDI, color = condition)) +
  geom_point(size = 2.5, alpha = 0.7) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.15, linewidth = 1) +
  scale_color_manual(values = c("Item Only" = "#9B59B6", 
                                 "Task Only" = "#E67E22", 
                                 "Both" = "#1ABC9C")) +
  labs(title = "Figure 11: Relationship Between Hit Rate and Lure Discrimination (LDI)",
       x = "Hit Rate (Target Recognition)", y = "Lure Discrimination Index (LDI)",
       color = "Condition") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", size = 14),
        legend.position = "bottom")

ggsave("figures/fig11_hit_vs_ldi.png", width = 10, height = 5, dpi = 300)
```

# 10. Summary of Key Findings

```{r summary}
cat("====================================================\n")
cat("       SUMMARY OF DESCRIPTIVE STATISTICS             \n")
cat("====================================================\n\n")

for (cond in levels(participant_metrics$condition)) {
  d <- participant_metrics[participant_metrics$condition == cond, ]
  cat(paste0("--- ", cond, " Condition ---\n"))
  cat(paste0("  N participants: ", nrow(d), "\n"))
  cat(paste0("  Hit Rate: ", round(mean(d$hit_rate, na.rm=T), 3), 
             " ± ", round(sd(d$hit_rate, na.rm=T), 3), "\n"))
  cat(paste0("  FA Rate:  ", round(mean(d$fa_rate, na.rm=T), 3), 
             " ± ", round(sd(d$fa_rate, na.rm=T), 3), "\n"))
  cat(paste0("  LDI:      ", round(mean(d$LDI, na.rm=T), 3), 
             " ± ", round(sd(d$LDI, na.rm=T), 3), "\n"))
  cat(paste0("  REC:      ", round(mean(d$REC, na.rm=T), 3), 
             " ± ", round(sd(d$REC, na.rm=T), 3), "\n"))
  cat(paste0("  Mean RT:  ", round(mean(d$rt_overall, na.rm=T), 3), 
             " ± ", round(sd(d$rt_overall, na.rm=T), 3), " sec\n\n"))
}
```

```{r session-info}
sessionInfo()
```
